False Positive Probabilities for all Kepler Objects of Interest: \\ \nvalnew\ newly validated planets and \nfpnew\ likely false positives

### 2015-12-09 ms.tex
new validation table written

\newcommand{\nbadphot}{28} %EmptyPhotometry or BadPhotometry of MissingKOI or IOError
%\newcommand{\nbadphotFP}{34}
%\newcommand{\nbadphotCAND}{13}
\newcommand{\nbadrowemcmc}{233}
\newcommand{\nbadstellar}{74} %MissingStellar
\newcommand{\nbadsec}{8} %NoWeakSecondary
%\newcommand{\nattempted}{8141}
%\newcommand{\ntryfail}{112}
\newcommand{\nbadtrapfit}{38}
\newcommand{\nbadroche}{37}
%\newcommand{\nbadother}{2}
%\newcommand{\nbadotherFP}{1}
%\newcommand{\nbadotherCAND}{1}
%\newcommand{\nbadotherCONF}{1}

\newcommand{\nreliable}{2943}
\newcommand{\nval}{2046}
\newcommand{\nreliableFP}{419}
\newcommand{\nvalnew}{1380}
\newcommand{\nfpnew}{424}
\newcommand{\ncurval}{1002}

@@

\shorttitle{False Positive Probabilities for KOIs}
\shortauthors{Morton et al.}

@@

This work presents results of applying \vespa\ \emph{en masse} to the
entire \kepler\ catalog. This is both the first time that most
\kepler\ candidates have been individually analyzed to assess false
positive probability and the first time that a detailed automated
planet validation calculation has been applied on such a large scale.
\Sect{methods} describes the methods used, \sect{results} presents and
discusses the results, and \sect{conclusions} contains concluding
remarks.

@@

To summarize, the results presented in \tab{fpp} are strictly reliable
only for KOIs \emph{that have already strongly passed the other Kepler
  vetting tests, and that are not indicated to be clearly poor fits to
  all the proposed hypotheses}.  The first cut for this is the KOI
disposition: FALSE POSITIVE indicates failure of one or more of these
tests (and thus probable invalidity of the \vespa\ calculation).
However, because of the generally permissive philosophy of the
\kepler\ dispositioning, not all CANDIDATE KOIs receive the same level
of vetting.  That is, something is not identified as a FALSE POSITIVE
\emph{unless there is positive confirmation of false positive status}.
In particular, when pixel-level analysis fails to determine whether
the signal is indeed coming from the target star because of low SNR,
such a KOI will still receive CANDIDATE disposition.  Therefore, the
\vespa-calculated FPP may be considered most reliable only when a KOI
is designated a CANDIDATE (or CONFIRMED) and has large enough SNR to
enable secure positional determination.

To enable interpretation, \tab{fpp} thus contains both the current KOI
disposition and the results of the positional probability calculations
of \citet{Bryson:KSCI}.  \Fig{fppall} shows FPPs for all KOIs passing
the following criteria:
\begin{itemize}
\item Dispositioned CANDIDATE or CONFIRMED at the NExScI archive,
\item $> 0.99$ probability to be on the target star, according
  to \citet{Bryson:KSCI}, along with a positional probability
  ``score'' $> \posprobthresh$ (indicating a reliable result), and   
\item $L_{\rm tot} > 10^{-3}$, indicating that not all the hypotheses are
  poor fits to the transit signal.
\end{itemize}

These selections leave \nreliable\ KOIs for which the \vespa\ results
can be considered most reliable.  Of these, \nval\ have ${\rm FPP} <
0.01$, and are thus validated at the 99\% level.  These are not all
new validations, however, since a number of these are already
CONFIRMED.  \Fig{fpppie} shows a different kind of summary, grouping
KOIs by disposition and splitting up the CANDIDATES according to the
\vespa\ results, showing that \nvalnew\ KOIs are newly validated at
the 99\% level.  \Fig{rpcand} shows the radii and periods of the
CANDIDATE KOIs, grouped by FPP.

One final note about these calculations is that unidentified transit
timing variations (TTVs) will increase the FPP of a transit signal, as
the shape of the folded light curve will be distorted, typically
resulting in a longer signal which a trapezoid model will also
identify as more V-shaped.  While we have analyzed light curves
correcting for known TTVs when available, we have also undoubtedly
missed many systems with as-yet-unspecified TTV signals.  This means
that a large FPP may be simply an indication of unidentified TTVs
rather than an astrophysical false positive, especially in
multi-planet systems, which should overall have a very low false
positive rate \citep{Lissauer:2014,Rowe:2014}.  \Tab{fpp} lists
whether known TTV were accounted for when constructing the folded
transit light curve for each KOI.

@@

We were not able to successfully run \vespa\ on all of the KOIs. The 
various reasons for these failures are detailed below.  
\begin{enumerate}
\item \nbadphot\ KOIs did not receive MCMC modeling.  Most of these
  have been are already designated FALSE POSITIVE at the archive.
\item \nbadrowemcmc\ KOIs did receive MCMC modeling but had unphysical fit
  results; e.g., negative $R_p/R_\star$ or best-fit impact parameter
  greater than $(1 + R_p/R_\star)$.  These KOIs were left out of the \vespa\ calculations.
\item The host stars of \nbadstellar\ KOIs are not included in the
  \citetalias{Huber:2014} stellar property catalog, and thus were not
  included in this analysis.
\item \nbadsec\ KOIs have no \verb|koi_depth_err1| value on the archive,
  and thus had no weak secondary constraint and were left out of the
  calculations.
\item For \nbadtrapfit\ KOIs, the trapezoid MCMC fit did not converge.  The
  convergence criterion was for the autocorrelation time of the chain
  for each parameter to be shorter than 10\% of the total chain
  length.
\item For \nbadroche\ KOIs, the orbital period and stellar properties of the 
  candidate imply the orbit to be within its host star's Roche limit.  This
  usually happens when the host star is estimated to be a giant, and
  these situations are nearly always false positives. 
%\item Finally, a small number of KOIs failed for as-yet-undetermined
%  reasons.  %, one of which is a FALSE POSITIVE (371.02) and one is a 
%  candidate (2433.06).
\end{enumerate}
The numbers in this list correspond to the numbers in
the ``failure'' column in \tab{fpp}.

@@

@@

These calculations depend largely on the results of
\citet{Bryson:KSCI}, which quantify the probability that the
eclipse/transit signal is spatially coincident on the sky with the
presumed target star.  Without confirmation that the transit signal is
not coming from a significantly displaced source, the sky area used as
part of the prior for the false positive scenarios would need to be
significantly larger than the positional uncertainty value assumed in
this work (described in \sect{methods:fpp}).  Additionally, the
blended false positive scenarios \vespa\ considers are assumed to be
fully contained within the photometric aperture; this assumption would
also be broken if the source of the eclipse were significantly
displaced.

>>>>

\newcommand{\nbadphot}{28} %EmptyPhotometry or BadPhotometry of MissingKOI or IOError
%\newcommand{\nbadphotFP}{34}
%\newcommand{\nbadphotCAND}{13}
\newcommand{\nbadrowemcmc}{233}
\newcommand{\nbadstellar}{74} %MissingStellar
\newcommand{\nbadsec}{8} %NoWeakSecondary
%\newcommand{\nattempted}{8141}
%\newcommand{\ntryfail}{112}
\newcommand{\nbadtrapfit}{38} %MCMCError or NoFitError
\newcommand{\nbadroche}{37}
%\newcommand{\nbadother}{2}
%\newcommand{\nbadotherFP}{1}
%\newcommand{\nbadotherCAND}{1}
%\newcommand{\nbadotherCONF}{1}

\newcommand{\nreliable}{2823} %from figure
\newcommand{\nval}{1994} %from figure
\newcommand{\nreliableFP}{398} %from figure
\newcommand{\nvalnew}{1324} %from pie chart
\newcommand{\nfpnew}{424}  % all candidate with FPP < 0.9
\newcommand{\ncurval}{984} % from pie chart

@@

\shorttitle{False Positive Probabilities for all KOIs}
\shortauthors{Morton et al.}

@@

This work presents results of applying \vespa\ \emph{en masse} to the
entire \kepler\ catalog. This is both the first time that most
\kepler\ candidates have been individually analyzed to assess false
positive probability and the first time that a detailed automated
planet validation calculation has been applied on such a large scale.
\Sect{methods} describes the methods used, \sect{data} describes the data set,
\sect{results} presents the results, \sect{comparisons} compares these
results with observational studies, and \sect{conclusions} contains concluding
remarks.

@@

\begin{deluxetable}{ccc}
\tabletypesize{\scriptsize}
\tablecaption{Mean FPPs of candidate KOIs with reliable \vespa\ calculations
  \tablabel{meanFPP}}
\tablehead{\colhead{Selection} & \colhead{Number} & \colhead{mean FPP}}
\startdata
all & 2823 & 0.148 \\
singles & 1665 & 0.196 \\
multis & 1158 & 0.078 \\
$R_p > 10\,R_\oplus$ & 347 & 0.704 \\ 
$4\,R_\oplus < R_p < 10\,R_\oplus$ & 245 & 0.182 \\ 
$2\,R_\oplus < R_p < 4\,R_\oplus$ & 1150 & 0.057 \\ 
$R_p < 2\,R_\oplus$ & 1081 & 0.058 
\enddata
\end{deluxetable}

To summarize, the results presented in \tab{fpp} are strictly reliable
only for KOIs \emph{that have already strongly passed the other Kepler
vetting tests, and that are not indicated to be clearly poor fits to
all the proposed hypotheses}.  The first cut for this is the KOI
disposition: FALSE POSITIVE indicates failure of one or more of these
tests (and thus probable invalidity of the \vespa\ calculation).
However, because of the generally permissive philosophy of the
\kepler\ dispositioning, not all CANDIDATE KOIs receive the same level
of vetting.  That is, something is not identified as a FALSE POSITIVE
\emph{unless there is positive confirmation of false positive status}.
In particular, when pixel-level analysis fails to determine whether
the signal is indeed coming from the target star because of low SNR,
such a KOI will still receive CANDIDATE disposition.  Therefore, the
\vespa-calculated FPP may be considered most reliable only when a KOI
is designated a CANDIDATE (or CONFIRMED) and has large enough SNR to
enable secure positional determination.  In addition, for greatest
reliability we also require a signal's multiple-event statistic (MES;
equivalend to SNR) to be greater than 10, in order to avoid low-SNR
signals that might be caused by light-curve systematics.

To enable interpretation, \tab{fpp} thus contains the current KOI
disposition, MES, and the results of the positional probability calculations
of \citet{Bryson:KSCI}.  \Fig{fppall} shows FPPs for all KOIs passing
the following criteria:
\begin{itemize}
\item Dispositioned CANDIDATE or CONFIRMED at the NExScI archive,
\item MES $>$ 10, indicating the signal is unlikely to be caused by 
      systematic noise in the light curve,
\item Probability $> 0.99$ to be on the target star, according
  to \citet{Bryson:KSCI}, along with a positional probability
  ``score'' $> \posprobthresh$ (indicating a reliable result), and   
\item $L_{\rm tot} > 10^{-3}$, indicating that not all the hypotheses are
  poor fits to the transit signal.
\end{itemize}

These selections leave \nreliable\ KOIs for which the \vespa\ results
can be considered most reliable;  \Tab{meanFPP} presents the mean FPPs
for different subsets, showing several notable features.  Single KOIs
are about 3$\times$ more likely to be false positives than KOIs in
multiple-KOI systems, in qualitative agreement with
\citet{Lissauer:2012}---and this is true even without giving any
``multiplicity boost'' to multi-KOI systems for the increased transit
probability of subsequent planets once one planet transits in a
coplanar system.  Also, large candidates typically have high FPPs, in
agreement with \citet{Santerne:2012} and \citet{Santerne:2015}.  And
finally, the mean FPPs are very consistent with the \textit{a priori}
predictions of \citet{Morton:2011b} and with the analysis of
\citet{Fressin:2013}.

Of these KOIs with reliable \vespa\ FPPs, \nval\ have ${\rm FPP}
<0.01$, and are thus validated at the 99\% level.  These are not all
new validations, however, since a number of these are already
CONFIRMED.  \Fig{fpppie} shows a different kind of summary, grouping
KOIs by disposition and splitting up the CANDIDATES according to the
\vespa\ results, showing that \nvalnew\ KOIs are newly validated at
the 99\% level. \Fig{rpcand} shows the radii and periods of the
CANDIDATE KOIs, grouped by FPP.

In addition to the confident validations, we identify \nfpnew\ KOIs
that currently have CANDIDATE disposition but for which \vespa\
calculates FPP $>$ 0.9; these KOIs are likely to be false positives.
We note that these likely false positives are not required to obey the
same selections that were done in order to ensure a clean sample for
validation.  This is because even if a signal has characteristics
such that a low \vespa\ FPP would not be  sufficient to validate it, a
high \vespa\ FPP is still sufficient to cast doubt on its planetary
nature.  As a demonstration of the ability of \vespa\ to identify 
false positives, \Fig{fppfps} shows the FPPs for KOIs that are 
dispositioned FALSE POSITIVE and have MES $>$ 10 and $L_{\rm tot} > 10^{-3}$.
The vast majority of these have large FPPs.

One final note about these calculations is that unidentified transit
timing variations (TTVs) will increase the FPP of a transit signal, as
the shape of the folded light curve will be distorted, typically
resulting in a longer signal which a trapezoid model will also
identify as more V-shaped.  While we have analyzed light curves
correcting for known TTVs when available, we have also undoubtedly
missed many systems with as-yet-unspecified TTV signals.  This means
that a large FPP may be simply an indication of unidentified TTVs
rather than an astrophysical false positive, especially in  
multi-planet systems, which should overall have a very low false positive
rate \citep{Lissauer:2014,Rowe:2014}.  It is thus probable that
despite multi-KOIs having lower FPPs than singles (\tab{meanFPP}),
even these relatively low FPPs are inflated by the effect of TTVs.
\Tab{fpp} lists whether known TTV were accounted for when constructing
the folded transit light curve for each KOI.

@@

We were not able to successfully run \vespa\ on all of the KOIs. The 
various reasons for these occasional failures are detailed below.  
\begin{enumerate}
\item \nbadphot\ KOIs did not receive MCMC modeling.  Most of these
  have been are already designated FALSE POSITIVE at the archive.
\item \nbadrowemcmc\ KOIs did receive MCMC modeling but had unphysical fit
  results; e.g., negative $R_p/R_\star$ or best-fit impact parameter
  greater than $(1 + R_p/R_\star)$.  These KOIs were left out of the \vespa\ calculations.
\item The host stars of \nbadstellar\ KOIs are not included in the
  \citetalias{Huber:2014} stellar property catalog, and thus were not
  included in this analysis.
\item \nbadsec\ KOIs have no \verb|koi_depth_err1| value on the archive,
  and thus had no weak secondary constraint and were left out of the
  calculations.
\item For \nbadtrapfit\ KOIs, the trapezoid MCMC fit did not converge.  The
  convergence criterion was for the autocorrelation time of the chain
  for each parameter to be shorter than 10\% of the total chain
  length.
\item For \nbadroche\ KOIs, the orbital period and stellar properties of the 
  candidate imply the orbit to be within its host star's Roche limit.  This
  usually happens when the host star is estimated to be a giant, and
  these situations are nearly always false positives. 
%\item Finally, a small number of KOIs failed for as-yet-undetermined
%  reasons.  %, one of which is a FALSE POSITIVE (371.02) and one is a 
%  candidate (2433.06).
\end{enumerate}
The numbers in this list correspond to the numbers in
the ``failure'' column in \tab{fpp}.

@@

We also briefly discuss another case of individual interest.
KOI-6705.01, a 0.99\,d signal around a mid M-dwarf star, was
identified  as a KOI of possible interest by \citet{Gaidos:2015}.
After significant follow-up observations and detailed full-frame image
analysis, they concluded that the signal was most likely due to a
charge-transfer  effect from a 1.99\,d EB located on the same CCD
column.  For this KOI, \vespa\ calculates an FPP of 1, with
probabilities of 0.86 to be a BEB, 0.13 to be a BEB with twice the
period (the true scenario uncovered by the detective work of
\citet{Gaidos:2015}), and 0.01 to be an HEB.

@@

The reliability of these calculations depends significantly on the
results of \citet{Bryson:KSCI}, which quantify the probability that
the eclipse/transit signal is spatially coincident on the sky with the
presumed target star.  Without confirmation that the transit signal is
not coming from a significantly displaced source, the sky area used as
part of the prior for the false positive scenarios would need to be
significantly larger than the positional uncertainty value assumed in
this work (described in \sect{methods:fpp}).  Additionally, the
blended false positive scenarios \vespa\ considers are assumed to be
fully contained within the photometric aperture; this assumption would
also be broken if the source of the eclipse were significantly
displaced.
